---
title: "Data Analysis"
author: "Katerina, Remy and Helena"
date: "11/15/2019"
output: pdf_document 
header-includes:
   - \usepackage{booktabs}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Add 1 paragraph describing your data set and 1 paragraph outlining your proposed in-class and out-of-class methods.

The Dialect Survey uses a series of questions, including rhyming word pairs and vocabulary words, to explore the distribution of dialects in American. We have 122 survey responses from 47,472 people from different city, state and zip code areas. The majority of participants were from the east coast, and approximately a third of the particpants were in the 20-29 age range.

In this project, we would use these data to fit a KNN classification model showing predicted class membership at each location in the US based on dialect. Our explanatory variable is zipcode and response is dialect. In addition, our second analysis will be based on a topic we haven't learned in class such as, neural networks to help us cluster, classify and recognize patterns of the data.  

```{r}
##Libraries
library(ggplot2)
library(dplyr)
library(readr)
library(caret)
dialect_survey<-read_csv("dialect_survey.csv")
zip_codes<-read_csv("zipcodes.csv")

##dataset with only ID, state, city, zip code, and answer to lawyer question
lawyer_response<-dialect_survey[,c(1:4, 18)]

for(i in seq_len(ncol(lawyer_response))){
  print(names(lawyer_response)[i])
  print(sum(is.na(lawyer_response[[i]])))
}

##we examined whether there are missing values and found 3 pieces of missing data for state, and 537 for city. However, there are no missing data values for zip code. We will use zip code as our explanatory variable.

##cleaned up our data: got rid of city and ID
lawyer_response<-dialect_survey[,c(3, 4, 18)]

lawyer_response_cleaned<-mutate(lawyer_response, ZIP=substr(ZIP, 2, 6))
lawyer_response_cleaned

results<-left_join(x=lawyer_response_cleaned, y=zip_codes, by="ZIP", copy=FALSE)
results<-results[!is.na(results$LAT) & !is.na(results$LNG), ]




# Get the world polygon and extract USA
library(maps)
USA <- map_data("world") %>% filter(region=="USA")

results$Q014 = factor(results$Q014,
                      levels = c("0","1","2","3","4"), labels=c("did not respond", "pronounced like 'boy' ", "pronounced like 'saw'", "use both interchangeably", "I don't know"))

# Left chart
g<-ggplot() +
  geom_polygon(data = USA, aes(x=long, y = lat, group = group), fill="grey") +
  geom_point(data=results, aes(x=LNG, y=LAT, color=Q014), size=0.7) + 
  theme_void()+coord_map(xlim = c(-180, -50),ylim = c(18, 72))
g

results %>% group_by(Q014)%>% count()


#KNN Classification Model

library(ISLR)

results$Q014 = factor(results$Q014,
                      levels = c("0","1","2","3","4"))

set.seed(87053)
train_inds <- caret::createDataPartition(results$Q014, p = 0.8)
Data_train <- results %>% dplyr::slice(train_inds[[1]])
Data_test <- results %>% dplyr::slice(-train_inds[[1]])
val_folds <- caret::createFolds(Data_train$Q014, k = 10)

#Select K for K nearest neighbors classification 

k_vals <- c(1:10, 25, 50, 75, 100, 150, 200, 250, 300)
results2 <- expand.grid(
    fold_ind = seq_len(10),
    k = k_vals,
    val_class_error = NA
  )
for(i in seq_len(10)) {
  train_data <- Data_train %>% dplyr::slice(-val_folds[[i]])
  val_data <- Data_train %>% dplyr::slice(val_folds[[i]])
  
  for(k in k_vals) {
    knn_fit <- train(
      form = Q014 ~LAT+LNG,
      data = train_data,
      method = "knn",
      preProcess = "scale",
      trControl = trainControl(method = "none"),
      tuneGrid = data.frame(k = k)
    )
    
    # get predicted values
    y_hats <- predict(knn_fit, newdata = val_data, type = "raw")
  
    # classification error rate
    save_ind <- which(results2$fold_ind == i & results2$k == k)
    results2$val_class_error[save_ind] <- mean(y_hats != val_data$Q014)
  }
}
results2 %>%
  group_by(k) %>%
  summarize(mean(val_class_error))

##Neural Network
library(neuralnet)

results_tiny_train<-results[1:10,]
results_tiny_test<-results[11:20,]


nn=neuralnet(Q014~LAT+LNG,data=results_tiny_train, hidden=3,act.fct = "logistic",
                linear.output = FALSE)
plot(nn)


predict_nn=compute(nn, results_tiny_test)

predict_nn$net.result


nn2=neuralnet(Q014~LAT+LNG,data=Data_train, hidden=3,act.fct = "logistic",
                linear.output = FALSE)
plot(nn2)

predict_nn2=compute(nn2, Data_test)

predict_nn2$net.result

pred<-predict(nn2, Data_test)

nn_classification_error<- table(Data_test$Q014, apply(pred, 1, which.max))





```

##https://www.r-graph-gallery.com/330-bubble-map-with-ggplot2.html
https://www.datacamp.com/community/tutorials/neural-network-models-r
